---
title: "Analysis of LCWMD 'Sonde'Contiunous' Data"
output: html_notebook
---
# Import LIbraries  
```{r}
library(tidyverse)
library(readr)
library(vcd)      # contains contingency table plotting fxns
library(emmeans)  # Provides tools for calculating marginal means
```


# Import Data
```{r}
sibfldnm    <- 'Derived Data'
parent      <- dirname(getwd())
sibling     <- file.path(parent,sibfldnm)

fn <- "Daily Data.csv"
fpath <- file.path(sibling, fn)

daily_medians <- read_csv(fpath, progress=FALSE) %>%
  select(-X1) %>%
  select(-ends_with('Min'), -ends_with('Max'), -ends_with('Mean'),
         -ends_with('Iqr'), -ends_with('SD'), -ends_with('_n'))

fn <- "Exceeds Data.csv"
fpath <- file.path(sibling, fn)
exceeds = read_csv(fpath, progress=FALSE) %>%
  select(-X1)

#summary(daily_data)
# rm(test.data)
```
  
```{r}
mosaic(~MaxT + ClassCDO, data = exceeds, color=TRUE)
```
So, that shows the two types of problems are highly correlated....  It may even be betterto treat temperatures as causal variables for DO in some kinds of models.


# Initial Models
```{r}
the_lm <- lm(log(Chl_Median) ~ factor(Month) + Site + Year + log(Precip+1) + log(PPrecip+1), data = daily_medians)
anova(the_lm)
```


```{r}
plot(the_lm)
```

No major problems with that model, although a few major outliers, with low leverage.  If anything the model is slightly light tailed, with a slight tendency to curvature, which suggests an underfit model. All terms are statistically significant, which is interesting, because I fit year as a linear term only. 

```{r}
old <- options() 
options(digits=3,scipen=2)
summary(the_lm)
options(old)
rm(old)
```
So:
1. Present-day precipitation supresses chlorides, but effect is much smaller than differences among sites.
2. Previous day precipitation ALSO suppresses chloride levels, but not by as much.
3. Chlorides have been INCREASING over the period of record
4. Site S01 and S017 have similar Chloride levels.
5.The highest chlorides are at site S03, followed by S01 and S17, S07, S05, and S06B.
6. The highest chlorides are in the winter months, especially February and March.
7. Chlorides are lowest in April, climbing again over the summer, and dropping in the fall.

I'd like to extract fits for each of these terms in meaningful ways.  I need adjusted or LS means....
Rather than fitting SITE as an independnet factor, it might be interesting to fit cummulative percent imperviousness, but that would take anested analysis.

```{r}
(month.emm <- emmeans(the_lm, ~Month, transform='response'))
```
```{r}
plot(month.emm) +
  ylab("Month")+
  xlab('Chlorides (mg/l)') +
  ggtitle('Marginal Geometric Means of Daily Medians') +
  theme_minimal() +
  coord_flip()
```

The following needs to be reordered to make it easier to interpret, so we need to chose an ordering of factors.
```{r}
site.emm <- emmeans(the_lm, ~Site, transform='response')
plot(site.emm) +
  ylab("Site")+
  xlab('Chlorides (mg/l)') +
  ggtitle('Marginal Geometric Means of Daily Medians') +
  theme_minimal() +
  coord_flip() 
```

```{r}
(i <- emmeans(the_lm, ~1))
```

```{r}
p <- exp(5.1 + (2010:2019 - 2014)*0.046983 )
dates = as.Date(paste(as.character(2010:2019), '06', '01', sep="-"))
df <- tibble(dates = dates, p=p)
plt <- ggplot() + geom_point(data = daily_medians, aes(x=sdate, y=Chl_Median, color=Site)) +
  geom_line(data =df, aes(x=dates, y=p), lwd = 2, color='blue') +
  xlab("Date")+
  ylab('Chlorides (mg/l)') +
  ggtitle('Daily Medians') +
  theme_minimal() +
  #scale_y_log10(limits = c(10,2000)) +
  geom_hline(yintercept=860, color='red') +
  geom_hline(yintercept=230, color='orange') 
plt
```

Note that because some sites were not sampled in earlier years, this slope may be misleading.

SO, it is easy to see that SEASONAL patterns and differences among SITES are much mor important than changes over time in salinity. 
